---
title: "Project Report"
output: 
  html_document: 
    toc: true
    toc_float: true
    code_folding: hide
---

```{r,echo=FALSE,warning=FALSE,message=FALSE}
library(tidyverse)
library(httr)
```

# Collaborators:
  •	Miriam Lachs ml5222
  <br /> •	Flora Pang fp2513
  <br /> •	Mari Sanders mos2128
  <br /> •	Lizy Choi jc6452
  
  
# Project Structure 

Based on the way we split up work we wanted to give a layout of how this project report is structured. The initial/general information about the project is given. Then each section details the individual analysis that was done. The final Appendices include the actual work/r code that was completed. 

1. General Project Information 
    - Research Questions
    - Project Motivation 
    - Project Significance 
    - Data used throughout
2. Activities
    - Data
    - EDA
    - Additional Analysis 
    - Discussion 
    - Shiny APP
3. Seasonal Regional Analysis 
4. Local NY 
5. Nature VS Machine 
6. Appendices/Work 

# General Information 
## Research Questions 

Our initial interest in national parks began with basic questions such as:  

- Are general historical trends reflected in national park data? 

- What are the seasonal visitation trends for national parks? 

- What kind of activities are offered by national parks? 

Our initial exploratory data analysis prompted us to ask more nuanced questions: 

- How can we wrangle complex and visualize traffic count data into an informative database for the wider knowledge community? 

- How can we categorize the nuances of seasonal trends in national parks? For example, seasonal trends based on activities, seasonal trends based on national park type.  

- How can we identify national parks that are decreasing in visitor interest? 

- Using a linear model, can an average visit to a national park be predicted by the number of activities offered by the park? 


## Project Motivation 

The project aims to investigate patterns and trends in visitation to national parks across the United States, with the central focus on understanding the various factors of influence such as visitation types, seasonal trends, historical events, and camping activities. As frequent enjoys of the parks, we hope to investigate what makes the parks popular and understand what tends to drive participation of the public. We hope that this investigation is useful for both people to get them into going to the parks and useful for park services to learn what increases participation. There are over 400 parks in the United States at this time. In 2023, there were 325.5 million recreation visits, an increase from 2022. Visitation at the parks has been increasing since the dip in visitation during 2020 and 2021, with closure of parks due to the COVID-19 Pandemic (NPS, 2024). We each chose an aspect of the investigation that appealed most to us. Therefore the project is split into different sections based on the area of interest for each collaborator. 

## Significance 

National Parks are an important part of American society, providing opportunities for recreation, education, and the preservation of historical landmarks, endangered landscapes, species, and ecosystems. These parks also play a significant role in boosting local economies by drawing visitors to nearby communities. However, accessing these parks often necessitates driving and parking vehicles, which can lead to increased human-wildlife interactions and the encroachment of parking lots on valuable park space. Such infrastructure disrupts the natural environment and diminishes the ecological integrity of these protected areas. 

Understanding when and why individuals visit National Parks offers valuable insights for fostering broader public engagement and enhancing visitor experiences. This paper examines the activities and parks that resonate most with visitors, providing a basis for strategic planning and resource allocation. Additionally, it explores the environmental impacts of vehicle usage in parks, highlighting opportunities for sustainable solutions that minimize harm to natural habitats. By identifying trends and preferences, these findings aim to promote greater park participation while ensuring the preservation of their natural and cultural heritage for future generations. 

## Data used

We started this project looking at two primary data sources. One looking at general Park information, from the NPS API. The second was a csv file generated from the NPS visitation query editor. 

```{r, echo=FALSE,message=FALSE,warning=FALSE}
visitation_data <- 
  read_csv("data/Query Builder for Public Use Statistics (1979 - Last Calendar Year).csv") %>% 
  janitor::clean_names() %>% 
  mutate(unit_code = tolower(unit_code)) %>% 
  rename(park_code = unit_code, 
         full_name = park_name) %>% 
  select(full_name, park_code, park_type, region, state, year, month, recreation_visits, tent_campers, rv_campers, tent_campers, backcountry)
```

Our visitation data includes variables `r names(visitation_data)`. It
has `r nrow(visitation_data)` rows and `r ncol(visitation_data)`
columns. The only data cleaning that needed to be done in the visitation
data set is to clean the column names and mutate some of the column names
to be able to merge the full national parks data with the visitation
data, as well as filtering to the data that we wanted to analyze.

-   `full_name`: Includes the full name of the National Park

-   `park_code`: Unique code that identifies each National Park

-   `park_type`: 18 Unique National Park Types

-   `state`: State(s) that the park is located

-   `year`: Year that the data is collected from

-   `month`: Month that the data is collected from

-   `recreation_visits`: Total number of recreation visits to each
    National Park

-   `tent_campers`: Number of visits that were tent campers

-   `rv_campers`: Number of visits that were RV campers

-   `backcountry`: Number of Backcountry hiker visits

```{r}
res=GET('https://developer.nps.gov/api/v1/parks?limit=500&api_key=B9nDpbkbrb3kSOjz6kXSxMJ3d6MSpUvt1QqYdeyn')

data = res %>% content("text") %>% jsonlite::fromJSON() %>% as_tibble()

NPS_data_parks=data %>% unnest(data) %>% select(fullName,parkCode,description,latitude,longitude,activities,states,entranceFees,operatingHours,designation) %>% unnest(activities,names_sep = '_') %>% unnest(entranceFees,names_sep = '_') %>% unnest(operatingHours,names_sep = '_') 
```

The Parks data came from the NPS API. To generate this data we first call the API, convert it from a JSON to a tibble, then select the information we are interested in and unnest the various data components. This cleaned data has the variables  `r names(NPS_data_parks)`. It has `r nrow(NPS_data_parks)` rows and `r ncol(NPS_data_parks)`
columns.

# Activities

## Data 

```{r}
res=GET('https://developer.nps.gov/api/v1/thingstodo?limit=5000&api_key=B9nDpbkbrb3kSOjz6kXSxMJ3d6MSpUvt1QqYdeyn')

things_data = res %>% content("text") %>% jsonlite::fromJSON() %>% as_tibble()

things_data=things_data %>% unnest(data) %>% unnest(relatedParks,names_sep = "_") %>% unnest(activities,names_sep = "_")

clean_things=things_data %>% select(title,relatedParks_parkCode,latitude,longitude,isReservationRequired,arePetsPermittedWithRestrictions,activities_id,activities_name,doFeesApply,arePetsPermitted,season,timeOfDay) %>%
  mutate(relatedParks_parkCode=str_to_upper(relatedParks_parkCode),
           latitude=as.numeric(latitude),
         longitude=as.numeric(longitude),
         isReservationRequired=as.logical(isReservationRequired),
         doFeesApply=as.logical(doFeesApply))
```

In addition to the general data used in this project, I also used another data set from the API. I also pulled the `thingstodo` data set. To clean this similarly to the parks data set, it needed to be unnested, relevant information selected and some data types converted. The final data set had the following variables `r names(clean_things)`. It has `r nrow(clean_things)` rows and `r ncol(clean_things)` columns.

## Exploratory Data Analysis

As exploratory data analysis, I looked into the activity counts across states and parks. I also looked at visitation trends. This is how I found I wanted to look into the relationship between activities and visitation. I also looked into entrance cost but ended up going in a different direction. However, it was after looking into the things to do data that I noticed there was a discrepancy in the initial parks data. Some activity categories were showing up under the things to do but not under the general park information. This means I had to go back and combine the data sets and redo the analysis to get a more accurate picture. 

## Analysis of Visitation and Activities

To do this analysis, after combining the data sets, I looked at the EDA graph showing the comparison of # of activities to average park visits per month. While slight I thought there may be a relationship and decided to run a linear regression.

## Discussion 

I found that the linear regression was significant. This meant there was a relationship between the number of activities offered by a park and the average number of visits to that park. Because of this positive relationship I wanted to look more into the advertised activities in the things to do data set. This is when I decided to build a shiny app to look into the different activities offered. 

## Shiny APP

The shiny app can be found [HERE](https://mlachs.shinyapps.io/shiny_app/). It allows you to find the various things to do based on activity category and other filters like reservations, pets and fees. 

## R code/ work 

My work was a mix of a working page [ml5222](https://mlachs.github.io/P8105_NPS_Final/ml5222.html) and the [Activities](https://mlachs.github.io/P8105_NPS_Final/Activities.html).

# Seasonal Regional Analysis 

## Exploratory Data Analysis

On initial analysis, it seems that July has the largest total visitation of any month. Therefore, we wanted to look at how different visitation looked like across seasons. Splitting the parks by month, there was clearly seasonal variation, with summer having the highest number of visitors across most park types. Winter recorded the lowest visitation levels, except for park types that specifically have activities that can be done during winter weather. More people visit national parks overall than any other park type, despite the fact that there are other park types with more parks across the U.S. There were some parks with variations, such as National Preserve, that had almost uniform visitation across the seasons. Looking at the activities offered there made it obvious that there were activities that could be done in all seasons.  

Analyzing activity-specific trends shows that tent camping is the most popular activity except in winter, where RV camping is the highest due to its ability to provide shelter in the winter weather. Backcountry hiking is the lowest in popularity across seasons, possibly due to the experience and training that it takes to complete a task such as this one. There were some park types that had none of these activities, including National Historic Site, National Memorial, National Battlefield, and International Historic Site. Additionally, there were parks that only have Backcountry visitation or RV camping.  Doing an analysis of regional trends on different activities offered in the parks shows that the west has the most visitation, especially in the summer months, despite the fact that the south has more visitation overall. This could be due to the more favorable weather conditions in the states that these parks are located that allow for more year-round visitation. 

## Discussion

Studying trends in seasonal and regional visitation shows that several things can be done to improve park visitation. If possible, parks with low winter visitation could benefit from the inclusion of winter-specific activities, such as snowshoeing. Additionally, improving facilities for RV and tent camping could help attract more visitors year-round. Using these results, the National Parks Service can make more data-informed decisions in order to optimize visitor experiences and ensure that the parks continue to preserve America’s precious landscapes. 

## Limitations

Many of the limitations in this project come from the data itself. One of the datasets was obtained through an API, where each of the columns was nested. In order to access any of the data contained in these columns would require several additional steps that is not usually needed. The two data sets used had different lengths, so joining them to use data in both was difficult Additionally, many of the parks had missing state data, which made it difficult to analyze regional trends. Further study could be conducted on more specific activities that are the most popular in the parks. The dataset obtained from an API included only the topics of the activities and not how many people do each activity. It would be interesting to do data analysis on these activities to see which ones are the most popular and could help improve park programming

## R Code 

My work is included in these two webpages:

[Regional Analysis](https://mlachs.github.io/P8105_NPS_Final/Regional_Analysis.html)

[Seasonal Analysis](https://mlachs.github.io/P8105_NPS_Final/Seasonal_Analysis.html)

# Local NY 

## Exploratory Data Analysis

The analysis reveals interesting trends across New York State parks. A key observation is the dominance of summer visitation, which aligns with warmer weather and increased outdoor activities. Parks such as Fire Island National Seashore attract the most visitors in summer, reflecting their appeal as seasonal destinations. Conversely, winter sees lower overall visitation except in parks equipped with activities suited for colder climates, such as Vanderbilt Mansion NHS, which exhibits more consistent attendance across all seasons.

Camping-specific trends highlight that tent camping remains the most popular activity except during winter, where RV camping surpasses it due to the shelter provided. Backcountry camping remains the least popular due to its challenging nature, requiring experience and preparation. Seasonal variations in visitation further suggest the impact of weather and regional characteristics on visitor behavior.

## Discussion

The findings underscore the importance of tailoring park activities and facilities to seasonal demands. For instance, parks with low winter visitation could consider offering winter-specific activities, such as skiing or snowshoeing, to attract more visitors. Enhancing camping facilities, including RV amenities and all-weather shelters, could encourage year-round visits. The consistent popularity of some parks across seasons, like Vanderbilt Mansion NHS, highlights the potential for expanding offerings to engage visitors beyond traditional peak seasons.

Moreover, understanding regional visitation trends can help the National Parks Service optimize resource allocation and marketing strategies. Promoting under-visited parks with unique activities could balance visitation across the state, ensuring all parks receive adequate attention and funding.

## Limitation

This analysis faced a few key challenges. First, the public use dataset required significant cleaning due to nested and inconsistent values, which may have affected the results. Missing regional data for some parks also made it harder to fully capture geographic trends. Additionally, the activity data only provided general topics, without details on participation rates, which limited insights into specific visitor preferences. Finally, trends like the 2020 visitation drop and the 2006 spring spike seem to align with events like the COVID-19 pandemic and unusual weather, but more data would be needed to confirm these patterns. Future studies addressing these gaps could provide richer, more actionable insights.

## R Code 

New York State Park Analysis can be found on this webpage. 

[Local NY](https://mlachs.github.io/P8105_NPS_Final/jc6452.html)


# Nature VS Machine 

## Data 

To do data analysis on national parks’ traffic count I used the Query Builder for National Park Traffic Counts.  

The datasets include data from 1985 to 2023.  

Provided that it was difficult to load data for every single national park into a single csv file, the National Park Traffic Count data was loaded separated by U.S. regions: firstly, North- East, secondly South-East, thirdly Alaska, Intermountain, and Mid-West regions, and lastly National-Capital and Pacific-West regions.  

For each region’s csv file included 15 variables: park name, unit code, park type, region, traffic counter, year, month, traffic count, park name total, unit code total, park type total, region total, traffic counter total, year total, month total, traffic counts total label, and traffic count total.  

National Park datasets included different traffic counters for the same national park. Where each counter would report a monthly count and also an annual count.  

## Exploratory Data Analysis and Discussion

In attempts to examine how our interactions with national parks have evolved over time and with no obvious topic consistently reported in the news, we determined that the most relevant development in recent history is the introduction of automobiles to the public. Many national parks emphasize preservation, but with increasing human contact, management needs to be flexible. To quantify how human-wildlife interactions have changed over the years, we can look at visitation pressure. High traffic volumes indicate heavy visitor use, leading to environmental degradation, pollution, and wildlife disturbance.  

Initial exploration of the data highlighted how traffic counts differ across types of national parks even within the same state. Furthermore, differ across the various traffic counters for each national park. These initial findings shifted focus towards annual mean traffic count for each state to provide a nationwide analysis of traffic impact on national parks, rather than in attempts to encompass every level of information.  

With the nuances of each state and traffic count data, we decided that rather than to create iterative analyses into each state and their national parks, we went down the path of curating and data wrangling to contribute towards the wider knowledge community through our shiny app.  

However, traffic counts may not be the best estimator for the impact of automobiles on human-wildlife interactions in national parks. This is because traffic count data does not specify the type of traffic it includes. It could be national park maintenance vehicles, visitors, or cars merely passing through without interacting with the park. To more directly quantify the impact of automobiles on national parks, examining parking lot usage may be more insightful. Parking lots are clearly man-made and reflect the park’s demand for car access, indicating efforts to accommodate visitors. By connecting traffic flow data with parking lot usage, we can better understand the needs of national parks and potentially improve traffic management and parking facilities. Plotly allows overlaying of two information datasets for a comprehensive view without cluttering the visual.  

## Shiny App

[This](https://fp2513.shinyapps.io/fp2513_shiny_dashboard/) Shiny app enables the selection of two different states: one for Chart 1 and another for Chart 2. This allows users to compare traffic counts side-by-side across years, with data further broken down by month. 

The app addresses previous limitations where we had to generalize data in our plots. By leveraging the app, users can gain a more detailed and interactive view of the traffic count dataset, enhancing the visualization experience and understanding of national park traffic counts. 

## Limitations

A key limitation of this analysis is that it represents only an initial step in the comprehensive study of traffic patterns and their implications in national parks. The current analysis, while insightful, provides a broad overview and suggests several avenues for further, more detailed research. Future studies should aim to delve deeper into specific aspects of human-wildlife interactions and environmental impacts.  

For instance, one critical area for further investigation is the quantification of pollution levels within national parks over the years. This could include analyzing the types and sources of pollutants, their temporal trends, and their effects on local ecosystems and wildlife.  

Additionally, understanding visitor experience and satisfaction through systematic surveys can offer valuable insights into the adequacy of current infrastructure. Surveys could assess the effectiveness of parking facilities, the sufficiency of amenities, and overall visitor satisfaction. This data would be instrumental in guiding resource allocation and infrastructure development to better meet visitor needs and protect park environments. This could involve spatial analysis to identify underserved areas and model the potential impact of proposed infrastructure improvements. 

## R Code 

The work on National Park Traffic Counts can be found on this webpage

[Nature vs Machines](https://mlachs.github.io/P8105_NPS_Final/fp2513.html)

# Final Remarks

This project underscored the importance of understanding trends and behaviors in national park visitation. Through the detailed analysis, we gained valuable insights into seasonal trends, park-specific visitation patterns, and even the broader environmental impact of human activity. The findings highlight a clear opportunity to tailor park offerings to visitor interests—such as introducing winter-specific activities or enhancing camping facilities to encourage year-round engagement. At the same time, the decline in interest for certain parks, like Vanderbilt Mansion NHS, signals a need for revitalized outreach and programming to sustain their cultural and historical significance. Moreover, addressing traffic impacts through improved data integration and infrastructure planning offers a pathway toward more sustainable park management. By embracing both visitor-focused and conservation-driven strategies, the National Parks Service can not only enhance public enjoyment but also protect these cherished landscapes for generations to come. This project serves as a foundation for future research, enabling deeper exploration of park trends, visitor preferences, and the balance between human presence and ecological preservation.
