---
title: "Regional Trends in Park Visitation"
output: 
  html_document: 
    toc: true
    toc_float: true
    code_folding: hide
---

```{r global options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r setup, include=FALSE}
library(tidyverse)
library(plotly)
library(httr)
library(leaflet)
library(patchwork)
knitr::opts_chunk$set(echo = TRUE, 
                      fig.width = 8, 
                      fig.height = 6,
                      out.width = "90%")
theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# **Data Importing and Cleaning**

```{r}
res = GET('https://developer.nps.gov/api/v1/parks?limit=500&api_key=B9nDpbkbrb3kSOjz6kXSxMJ3d6MSpUvt1QqYdeyn')

data = res %>% 
  content("text") %>% 
  jsonlite::fromJSON() %>% 
  as_tibble()

NPS_data = data %>% 
  unnest(data) %>% 
  select(fullName,latitude,longitude,topics, activities,states, parkCode) %>%  janitor::clean_names() %>% 
  mutate(
    latitude = as.numeric(latitude), 
    longitude = as.numeric(longitude)
  ) %>% 
  unnest(activities, names_sep = "_") %>% 
  unnest(topics, names_sep = "_")
visitation_data <- 
  read_csv("data/Query Builder for Public Use Statistics (1979 - Last Calendar Year).csv") %>% 
  janitor::clean_names() %>% 
  mutate(unit_code = tolower(unit_code)) %>% 
  rename(park_code = unit_code, 
         full_name = park_name) %>% 
  select(full_name, park_code, park_type, region, state, year, month, recreation_visits, tent_campers, rv_campers, tent_campers, backcountry)

combined_data <- full_join(NPS_data, visitation_data, by = c("park_code"))
```

# **Regional Comparisons**

```{r}
region_data_vists <- visitation_data %>% 
  mutate(region = case_when(
    state %in% c("CT", "RI", "NH", "VT", "NJ", "NY", "PA", "MD", "ME", "MA") ~ "northeast", 
    state %in% c("IL","IN", "MI", "OH", "WI", "IA", "KS", "MN", "MO", "NE", "ND", "SD") ~ "midwest", 
    state %in% c("FL", "GA", "NC", "SC", "VA", "DE", "WV", "AL", "KY", "MS", "TN", "AR", "LA", "OK", "TX", "DC") ~ "south", 
    state %in% c("AK", "CA", "HI", "OR", "WA", "AZ", "CO", "ID", "MT", "NV", "NM", "UT", "WY") ~ "west",
    state %in% c("VI", "AS", "GU", "PR") ~ "u.s. territory",
    TRUE ~ "no state data"
  ))
```

```{r}
region_data <- combined_data %>% 
  mutate(region = case_when(
    state %in% 
      c("CT", "RI", "NH", "VT", "NJ", "NY", "PA", "MD", "ME", "MA") ~ "northeast", 
    state %in% 
      c("IL","IN", "MI", "OH", "WI", "IA", "KS", "MN", "MO",
        "NE", "ND", "SD") ~ "midwest", 
    state %in% 
      c("FL", "GA", "NC", "SC", "VA", "DE", "WV", "AL", "KY", 
        "MS", "TN", "AR", "LA", "OK", "TX", "DC") ~ "south", 
    state %in% c("AK", "CA", "HI", "OR", "WA", 
                 "AZ", "CO", "ID", "MT", "NV", "NM", "UT", "WY") ~ "west",
    state %in% c("VI", "AS", "GU", "PR") ~ "u.s. territory",
    state == TRUE ~ "no state data"
  )) %>% select(c(-full_name.y, -topics_id, -topics_name, -activities_id))
```

## **Northeast**

```{r}
region_data %>%
  filter(region == "northeast") %>% 
  distinct(activities_name) %>% 
  head() %>% knitr::kable()

northeast_plot <- region_data %>%
  filter(region == "northeast") %>% 
  mutate(season = case_when(
      month %in% c(12, 1, 2) ~ "Winter", 
      month %in% c(3, 4, 5) ~ "Spring", 
      month %in% c(6, 7, 8) ~ "Summer", 
      TRUE ~ "Fall"
    )) %>% 
    group_by(season) %>% 
    summarize(
      mean_tent = mean(tent_campers, na.rm = TRUE), 
      mean_backcountry = mean(backcountry, na.rm = TRUE), 
      mean_rv = mean(rv_campers, na.rm = TRUE)
    ) %>% 
    pivot_longer(
      cols = starts_with("mean_"), 
      names_to = "type_visit", 
      values_to = "mean", 
      names_prefix = "mean_"
    ) %>% ggplot(aes(x = season, y = mean)) + 
  geom_col() +  
  facet_grid(~type_visit) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + 
  labs(title = "Visitation in Northeast by Season", 
       x = "Season", 
       y = "Average Visitation")
northeast_plot
```

## **Midwest**

```{r}
region_data %>%
  filter(region == "midwest") %>% distinct(activities_name) %>% 
  head() %>%  knitr::kable()

midwest_plot <- 
  region_data %>%
  filter(region == "midwest") %>% 
  mutate(season = case_when(
      month %in% c(12, 1, 2) ~ "Winter", 
      month %in% c(3, 4, 5) ~ "Spring", 
      month %in% c(6, 7, 8) ~ "Summer", 
      TRUE ~ "Fall"
    )) %>% 
    group_by(season) %>% 
    summarize(
      mean_tent = mean(tent_campers, na.rm = TRUE), 
      mean_backcountry = mean(backcountry, na.rm = TRUE), 
      mean_rv = mean(rv_campers, na.rm = TRUE)
    ) %>% 
    pivot_longer(
      cols = starts_with("mean_"), 
      names_to = "type_visit", 
      values_to = "mean", 
      names_prefix = "mean_"
    ) %>% ggplot(aes(x = season, y = mean)) + 
  geom_col() +  
  facet_grid(~type_visit) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + 
  labs(title = "Visitation in Midwest by Season", 
       x = "Season", 
       y = "Average Visitation")
midwest_plot
```

## **South**

```{r}
region_data %>%
  filter(region == "south") %>%
  distinct(activities_name) %>% 
  head() %>%  
  knitr::kable()

south_plot <- 
  region_data %>%
  filter(region == "south") %>% 
  mutate(season = case_when(
      month %in% c(12, 1, 2) ~ "Winter", 
      month %in% c(3, 4, 5) ~ "Spring", 
      month %in% c(6, 7, 8) ~ "Summer", 
      TRUE ~ "Fall"
    )) %>% 
    group_by(season) %>% 
    summarize(
      mean_tent = mean(tent_campers, na.rm = TRUE), 
      mean_backcountry = mean(backcountry, na.rm = TRUE), 
      mean_rv = mean(rv_campers, na.rm = TRUE)
    ) %>% 
    pivot_longer(
      cols = starts_with("mean_"), 
      names_to = "type_visit", 
      values_to = "mean", 
      names_prefix = "mean_"
    ) %>% ggplot(aes(x = season, y = mean)) + 
  geom_col() +  
  facet_grid(~type_visit) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + 
  labs(title = "Visitation South by Season", 
       x = "Season", 
       y = "Average Visitation")
south_plot
```

## **West**

```{r}
region_data %>% filter(region == "west") %>% distinct(activities_name) %>% 
  head() %>% knitr::kable()

west_plot <- 
  region_data %>% 
  filter(region == "west") %>% 
  mutate(season = case_when(
      month %in% c(12, 1, 2) ~ "Winter", 
      month %in% c(3, 4, 5) ~ "Spring", 
      month %in% c(6, 7, 8) ~ "Summer", 
      TRUE ~ "Fall"
    )) %>% 
    group_by(season) %>% 
    summarize(
      mean_tent = mean(tent_campers, na.rm = TRUE), 
      mean_backcountry = mean(backcountry, na.rm = TRUE), 
      mean_rv = mean(rv_campers, na.rm = TRUE)
    ) %>% 
    pivot_longer(
      cols = starts_with("mean_"), 
      names_to = "type_visit", 
      values_to = "mean", 
      names_prefix = "mean_"
    ) %>% ggplot(aes(x = season, y = mean)) + 
  geom_col() +  
  facet_grid(~type_visit) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + 
  labs(title = "Visitation in West by Season", 
       x = "Season", 
       y = "Average Visitation") 

northeast_plot + midwest_plot + south_plot + west_plot
```

Across all regions, it seems that summer has the highest visitation
across regions, with the exception of backcountry hiking, which has a
peak in spring.

For the Northeastern U.S. tent camping is by far the most popular
visitation type. Interestingly, in this region of the United States,
there was hardly any visitation in winter in all three visitation types.
This might be due to the fact that the Northeastern U.S. typically has
colder temperatures and possibly snow during this time.

For the Midwestern U.S., tent camping is the most popular visitation
type, especially in the summer, but it is closely followed by
backcountry hiking in summer. RV camping does seem to be higher in the
Midwest than in the northeastern region, especially during the summer.
As before, there is hardly any visitation in the winter, and actually
none for RV camping. This again might be due to weather conditions not
permitting outdoor activities such as these. For the Southern U.S., tent
camping has the highest average visitation in the summer time, closely
followed by RV camping. In the south, RV camping is much more popular
than in other regions around the U.S. Backcountry is lower, but does
have an almost even spread among the seasons. This could be due to the
more temperate conditions in the winter and other times of the year,
allowing individuals to enjoy this activity year round.

For the Western U.S., ten camping once again has the highest average
visitation across all seasons. Interestingly, backcountry hiking and RV
camping are equal in the summer time in this region. This region has
some visitation in the winter, but less than the southern region.

# **Comparing Regions**

```{r}
region_data %>% group_by(region) %>% 
  summarize(avg_visitation = mean(recreation_visits, na.rm = TRUE)) %>% 
  arrange(desc(avg_visitation)) %>% knitr::kable()

region_long <- region_data %>%
  pivot_longer(cols = c(recreation_visits, tent_campers, rv_campers, backcountry), 
               names_to = "visit_type", 
               values_to = "count") 
```

The western region of the U.S. seems to have the highest average
visitation of all regions, followed by the southern region.

Looking at the amount of parks in each region will be helpful to
determine if this is due to the fact that there are more parks in this
region, or if there is something else going on.

```{r}
region_data %>% 
  group_by(region) %>% 
  summarize(parks = n_distinct(park_code)) %>% knitr::kable()
```

From this table, we can see that actually the southern region has the
most amount of parks in this dataset, followed by the western region. So
this trend is visitation is not just due to the amount of parks in a
specific region.

```{r}
region_data %>%
  group_by(region) %>%
  summarize(
    mean_tent = mean(tent_campers, na.rm = TRUE),
    mean_backcountry = mean(backcountry, na.rm = TRUE),
    mean_rv = mean(rv_campers, na.rm = TRUE)
  ) %>%
  pivot_longer(
    cols = starts_with("mean_"),
    names_to = "type_visit",
    values_to = "mean",
    names_prefix = "mean_"
  ) %>% ggplot(aes(x = region, y = mean)) + 
  geom_col() + facet_grid(~type_visit) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

Now looking at the specific visits by region, we can see once again that
west has the highest mean visitation among all the regions, with its
highest being tent camping. Unsurprisingly, we see that the south has
the second highest mean visitation. Interestingly, in backcountry
hiking, the midwest has the second highest mean visitation, despite the
fact that it has one of the lowest amount of parks in the dataset.

## **US National Parks by Region**

```{r}
regional_data <- NPS_data %>% 
  mutate(region = case_when(
    states %in% c("CT", "RI", "NH", "VT", "NJ", "NY", 
                  "PA", "MD", "ME", "MA") ~ "Northeast", 
    states %in% c("IL","IN", "MI", "OH", "WI", "IA", "KS", 
                  "MN", "MO", "NE", "ND", "SD") ~ "Midwest", 
    states %in% c("FL", "GA", "NC", "SC", "VA", "DE", "WV", 
                  "AL", "KY", "MS", "TN", "AR", "LA", "OK", "TX", "DC") ~ "South", 
    states %in% c("AK", "CA", "HI", "OR", "WA", "AZ", 
                  "CO", "ID", "MT", "NV", "NM", "UT", "WY") ~ "West",
    states %in% c("VI", "AS", "GU", "PR") ~ "U.S. Territory", 
    length(unique(states)) > 1 ~ "Across Multiple States"
  )) 
g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showland = TRUE,
  landcolor = toRGB("#e5ecf6"),
  countrywidth = 0.5,
  subunitwidth = 0.5
)

fig <- plot_geo(regional_data, lat = ~latitude, lon = ~longitude)
fig <- fig %>% add_markers(
    text = ~paste(full_name, states, sep = "<br />"),
    color = ~region, symbol = I("circle"), size = I(8), hoverinfo = "text"
  )

fig <- fig %>% layout(
    title = 'US National Parks', geo = g
  )
fig
```

Finally, we wanted a way to visualize the national parks on a U.S. map.