---
title: "Seasonal Trends in Park Visitation"
output: 
  html_document: 
    toc: true
    toc_float: true
    code_folding: hide
---

```{r global options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r setup, include=FALSE}
library(tidyverse)
library(plotly)
library(httr)
library(leaflet)
library(patchwork)
knitr::opts_chunk$set(echo = TRUE, 
                      fig.width = 8, 
                      fig.height = 6,
                      out.width = "90%")
theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# **Data Importing and Cleaning**

```{r}
res = GET('https://developer.nps.gov/api/v1/parks?limit=500&api_key=B9nDpbkbrb3kSOjz6kXSxMJ3d6MSpUvt1QqYdeyn')

data = res %>% 
  content("text") %>% 
  jsonlite::fromJSON() %>% 
  as_tibble()

NPS_data = data %>% 
  unnest(data) %>% 
  select(fullName,latitude,longitude,topics, activities,states, parkCode) %>%  janitor::clean_names() %>% 
  mutate(
    latitude = as.numeric(latitude), 
    longitude = as.numeric(longitude)
  ) %>% 
  unnest(activities, names_sep = "_") %>% 
  unnest(topics, names_sep = "_")
visitation_data <- 
  read_csv("data/Query Builder for Public Use Statistics (1979 - Last Calendar Year).csv") %>% 
  janitor::clean_names() %>% 
  mutate(unit_code = tolower(unit_code)) %>% 
  rename(park_code = unit_code, 
         full_name = park_name) %>% 
  select(full_name, park_code, park_type, region, state, year, month, recreation_visits, tent_campers, rv_campers, tent_campers, backcountry)

combined_data <- full_join(NPS_data, visitation_data, by = c("park_code"))
```

# **Trends across Time**

## **Total Trend by Park Type Across Months**

```{r}
visitation_data %>% 
  group_by(park_type, month) %>% 
  summarize(total_visitation = sum(recreation_visits)) %>% 
plot_ly(x = ~as.factor(month), y = ~total_visitation, type = "scatter", 
        mode = "lines", color = ~park_type)

visitation_data %>% 
  group_by(park_type) %>% 
  summarize(parks = n_distinct(full_name)) %>%
  arrange(desc(parks))  %>%  
  knitr::kable()
```

As seen in the graph, national parks have the highest total visitation
among all park types. Interestingly, there more national monuments,
national historic sites, and national historic parks than national
parks, even though all of these have less total visitation than national
parks.

## **Total Visitation by Visitation Type and Season**

The `visitation_data` data set also includes information on visitation
split by `tent_campers`, `backcountry`, and `rv_campers`. We were
interested in looking at the total visitation split by these three
visitation types and by season.

```{r}
visitation_data %>% 
  mutate(season = case_when(
    month %in% c(12,1,2) ~ "Winter", 
    month %in% c(3,4,5) ~ "Spring", 
    month %in% c(6,7,8) ~ "Summer", 
    TRUE ~ "Fall"
  )) %>% 
  group_by(season) %>% 
  summarize(total_tent = sum(tent_campers), 
            backcountry_visits = sum(backcountry), 
            total_rv = sum(rv_campers)) %>% 
  pivot_longer(
    total_tent:total_rv, 
    values_to = "total_visit", 
    names_to = "type_visit",
    names_prefix = "total_"
  ) %>% 
  plot_ly(x = ~season, y = ~total_visit, color = ~type_visit, type = "bar")
```

In fall, winter, and spring, the highest visitation is `tent_campers`,
with a peak summer. Interestingly and unsurprisingly, the highest
visitation in winter is `rv_campers`. This is probably due to weather
conditions not permitting tent camping, but allows for rv camping. For
fall, winter, and spring there is a similar trend where `backcountry`
has the lowest total visitation and `tent_campers` has the highest
visitation. In winter, `rv_campers` has the highest total visitation,
then `tent_campers`, and then `backcountry`.

## **Average Park Visitation by Season and Park Type**

Next, we were interested in seeing the trends in park visitation by
season and park type.

```{r}
park_types <- unique(visitation_data$park_type)
plots <- list()
for (i in seq_along(park_types)) {
plots[[i]] <- visitation_data %>% 
    filter(park_type == park_types[i]) %>% 
    mutate(season = case_when(
      month %in% c(12, 1, 2) ~ "Winter", 
      month %in% c(3, 4, 5) ~ "Spring", 
      month %in% c(6, 7, 8) ~ "Summer", 
      TRUE ~ "Fall"
    )) %>% 
    group_by(season) %>%
    summarize(avg_visits = mean(recreation_visits, na.rm = TRUE)) %>%
    ggplot(aes(x = season, y = avg_visits, fill = season)) +
    geom_col() +
    ggtitle(paste( park_types[i])) 
} 
```

Most of the parks have a similar trend in average visitation, where
summer is the highest and winter is the lowest.

## **National Lakeshore Seasonal Trends and Activities**

```{r, message= FALSE}
plots[7]
```

For `national lakeshore` parks, there is a really stark difference in
average visits between summer and the other seasons.

```{r}
activities_count <- 
  combined_data %>% 
  filter(park_type == "National Lakeshore") %>% 
  distinct(activities_name) %>% 
  nrow()

combined_data %>% 
  filter(park_type == "National Lakeshore") %>% 
  distinct(activities_name) %>% 
  head() %>% knitr::kable()
```

Based on the activities offered in National Lakeshore, we can see that
most of the activities revolve around water, such as `boating`,
`sailing`, and `boat tours`. Interestingly, at National Lakeshore, there
are skiing activities, but the winter average visits are still the
lowest.

## **National Preserve Seasonal Trends and Activities**

```{r, message= FALSE}
plots[10]
```

Another interesting pattern is that in National Preserve, there is a
higher average visitation in spring than summer and winter has a higher
visitation than fall.

```{r}
activities_count <- 
  combined_data %>% 
  filter(park_type == "National Preserve") %>% 
  distinct(activities_name) %>% 
  nrow()
combined_data %>% 
  filter(park_type == "National Preserve") %>% 
  distinct(activities_name)  %>% 
  head() %>% 
  knitr::kable()
```

There are `r activities_count` activities offered at this park. It is
near the middle of the distribution of total activities, so the trend is
not due to having many more activities offered than other park types. It
seems like there is a much bigger variety in the types of activities
offered at these parks. There are tours, stargazing, biking, picnicking,
hunting, water activities, wildlife watching, museums and stores, as
well as winter activities like skiing, and snowshoeing.

## **International Historic Site Seasonal Trends and Activities**

```{r, message= FALSE}
plots[18]
```

This plot is interesting because there is no average visitation in
winter.

```{r}
activities_count <- 
  combined_data %>% 
  filter(park_type == "International Historic Site") %>% 
  distinct(activities_name) %>%
  nrow()
combined_data %>% 
  filter(park_type == "International Historic Site") %>% 
  distinct(activities_name) %>%
  head() %>% 
  knitr::kable()
```

There are `r activities_count` activities offered at this park type. It
makes sense that there is no visitation in the winter because the only
activities offered are `paddling`, `junior ranger program` and
`wildlife watching`. There are no activities that would really be done
in the winter at these parks.

## **Average Visitation by Visit and Park Type**

```{r, message = FALSE}
park_types <- unique(visitation_data$park_type)
plots <- list()

for (i in seq_along(park_types)) {
  visit_summary <- visitation_data %>% 
    drop_na() %>% 
    filter(park_type == park_types[i]) %>% 
    mutate(season = case_when(
      month %in% c(12, 1, 2) ~ "Winter", 
      month %in% c(3, 4, 5) ~ "Spring", 
      month %in% c(6, 7, 8) ~ "Summer", 
      TRUE ~ "Fall"
    )) %>% 
    group_by(season) %>% 
    summarize(
      mean_tent = mean(tent_campers, na.rm = TRUE), 
      mean_backcountry = mean(backcountry, na.rm = TRUE), 
      mean_rv = mean(rv_campers, na.rm = TRUE)
    ) %>% 
    pivot_longer(
      cols = starts_with("mean_"), 
      names_to = "type_visit", 
      values_to = "mean", 
      names_prefix = "mean_"
    ) 
  plots[[i]] <- ggplot(visit_summary, aes(x = season, y = mean)) + 
    geom_col() + 
    facet_grid(~type_visit) +
    ggtitle(paste(park_types[i])) + 
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
} 

```

## **No Visitation**

```{r, message = FALSE}
plots[[4]] + plots[[8]] + plots[[17]] + plots[[18]]
```

National Historic Sites, National Memorials, National Battlefield Parks,
International Historic Sites all had no faceted average visits. These
four park types do not have any `rv_campers`, `backcountry` or
`tent_campers`.

## **Only Backcountry Visitation**

```{r, message= FALSE}
plots[[16]] + plots[[15]] + plots[[6]]
```

`National Reserve` only had backcountry visitation in spring.
`National Military Park` had backcountry visitation in all seasons, with
mean visitation being highest in spring. `National Battlefield` had
backcountry visitation in all seasons, with mean visitation being lowest
in summer and highest in spring.

## **No RV Visitation**

```{r, message= FALSE}
plots[13]
```

National Wild and Scienic River has `backcountry` and `tent` visitation,
no `rv` visitation. Interestingly though, the mean `tent` visitation is
low for all seasons, with a slight peak in summer.

## **Park (Other) Visitation**

```{r, message= FALSE}
plots[14]
```

This plot was interesting because the park type `Park (Other)` has high
`tent` mean visitation, `rv` visitation, but low `backcountry`
visitation. It is the highest in spring and summer.

